---
title: "IA2030 vaccine impact results"
author: "Austin Carter"
date: "3/11/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(data.table); library(ggplot2); library(knitr); library(forcats)
```

```{r, echo = F}
dt <- fread("outputs/detailed_results.csv")
```

## Global annual total with baseline
```{r, results = 'asis', echo = F}
dt[, baseline := total - incremental]
global_baseline <- sum(dt[year == 2021]$baseline)
global_ann <- rbind(
  data.table(year = "2019 (baseline)", `Total deaths averted` = global_baseline),
  dt[, .(`Total deaths averted` = sum(total)), by = year]
)
setnames(global_ann, "year", "Year")
write.csv(global_ann, "results/global_ann.csv", row.names = F)
kable(global_ann)
```

## By disease global

```{r, echo = F}
disease_year_dt <- dt[, .(deaths_averted = sum(total, na.rm = T)), by = .(disease, year)] 
my_colors1 <- c(RColorBrewer::brewer.pal(name = "Paired", n = 12), c("darkblue", "darkgreen"))
my_colors1[1] <- "hotpink"
my_colors2<- rev(RColorBrewer::brewer.pal(name = "Paired", n = 3))
## By vaccine
gg <- ggplot(disease_year_dt[year %in% 2000:2030], aes(x = year, y = deaths_averted / 1e6, fill = disease)) +
    geom_bar(position = "stack", stat = "identity") +
    scale_fill_manual(values = my_colors1, name = "Disease") +
    theme_bw() + xlab("Year") + ylab("Deaths averted (in millions)") +
    ggtitle("Global deaths averted") +
  scale_x_continuous(breaks = scales::pretty_breaks())
gg
png("results/global_by_disease.png")
print(gg)
dev.off()
disease_year_dt[, deaths_averted := round(deaths_averted / 1e3)]
cast_dt <- dcast(disease_year_dt, disease ~ year, value.var = "deaths_averted")
kable(cast_dt)
cast_dt
write.csv(cast_dt, "results/disease_year.csv", row.names = F)
```

## By disease for each region

```{r, echo = F}
region_d_dt <- dt[, .(deaths_averted = sum(total)), by = .(region, year, disease)]
full_dt <- data.table(expand.grid(
    region = unique(region_d_dt$region),
    year = unique(region_d_dt$year),
    disease = unique(region_d_dt$disease)
))
region_d_dt <- merge(region_d_dt, full_dt, all.y = T)
region_d_dt[is.na(deaths_averted), deaths_averted := 0]

gg <- ggplot(region_d_dt[year %in% 2021:2030], aes(x = year, y = deaths_averted / 1e6 , fill = disease)) +
geom_bar(position = "stack", stat = "identity") +
scale_fill_manual(values = my_colors1, name = "Disease") +
theme_bw() + xlab("Year") + ylab("Deaths averted (in millions)") +
facet_wrap(~region) +
scale_x_continuous(breaks = scales::pretty_breaks()) +
scale_y_continuous(breaks = scales::pretty_breaks())
print(gg)
png(paste0("results/region_by_disease.png"))
print(gg)
dev.off()

```

## Total by year by region
```{r, echo = F}
region_y_dt <- dt[, .(deaths_averted = sum(total)), by = .(region, year)]
my_colors3 <- RColorBrewer::brewer.pal(name = "Paired", n = 6)

gg <- ggplot(region_y_dt[year %in% 2021:2030], aes(x = year, y = deaths_averted / 1e6 , fill = region)) +
  geom_area(color = "white", alpha = 0.8) +
  scale_fill_manual(values = my_colors3, name = "Region") +
  theme_bw() + xlab("Year") + ylab("Deaths averted (in millions)") +
  ggtitle("Deaths averted by region") +
  scale_x_continuous(breaks = scales::pretty_breaks())
print(gg)
png("results/total_by_region.png")
print(gg)
dev.off()
```

## By disease mean annual total deaths averted with min and max annual
```{r, echo = F}
ann_d_dt <- dt[, .(total = sum(total)), by = .(year, disease)]
d_summ <- ann_d_dt[, .(mean = mean(total), min = min(total), max = max(total)), by = disease]
ann_dt <- dt[, .(total = sum(total)), by = year]
t_summ <- ann_dt[, .(mean = mean(total), min = min(total), max = max(total))]
t_summ[, disease := "Total"]
d <- rbind(d_summ, t_summ)
gg <- ggplot(d_summ, aes(x = disease, y = mean / 1e3)) + geom_point() +
  geom_errorbar(aes(ymin = min / 1e3, ymax = max / 1e3), width = 0.2) +
  theme_bw() + xlab("Disease") + ylab("Annual total deaths averted (in thousands)") +
  ggtitle("Mean and range of annual total deaths averted by disease")
print(gg)
png("results/mean_range_disease.png")
print(gg)
dev.off()
kable(d)
```


# Incremental

## Global annual incremental
```{r, results = 'asis', echo = F}
global_ann <- dt[, .(`Incremental deaths averted` = sum(incremental)), by = year]
setnames(global_ann, "year", "Year")
kable(global_ann)
```

## By disease global

```{r, echo = F}
disease_year_dt <- dt[, .(deaths_averted = sum(incremental, na.rm = T)), by = .(disease, year)] 
my_colors2<- rev(RColorBrewer::brewer.pal(name = "Paired", n = 3))
## By vaccine
gg <- ggplot(disease_year_dt[year %in% 2000:2030], aes(x = year, y = deaths_averted / 1e6, fill = disease)) +
    geom_area(color = "white", alpha = 0.8) +
    scale_fill_manual(values = my_colors1, name = "Disease") +
    theme_bw() + xlab("Year") + ylab("Deaths averted (in millions by YoV)") +
    ggtitle("Global deaths averted") +
  scale_x_continuous(breaks = scales::pretty_breaks())
gg
png("results/inc_global_by_disease.png")
print(gg)
dev.off()
cast_dt <- dcast(disease_year_dt, disease ~ year, value.var = "deaths_averted")
kable(cast_dt)
```

## By disease for each region

```{r, echo = F}
region_d_dt <- dt[, .(deaths_averted = sum(incremental)), by = .(region, year, disease)]
full_dt <- data.table(expand.grid(
    region = unique(region_d_dt$region),
    year = unique(region_d_dt$year),
    disease = unique(region_d_dt$disease)
))
region_d_dt <- merge(region_d_dt, full_dt, all.y = T)
region_d_dt[is.na(deaths_averted), deaths_averted := 0]
for(r in unique(region_d_dt$region)) {
    plot_dt <- region_d_dt[region == r]
    gg <- ggplot(plot_dt[year %in% 2021:2030], aes(x = year, y = deaths_averted , fill = disease)) +
    geom_area(color = "white", alpha = 0.8) +
    scale_fill_manual(values = my_colors1, name = "Disease") +
    theme_bw() + xlab("Year") + ylab("Deaths averted") +
    ggtitle(r) +
  scale_x_continuous(breaks = scales::pretty_breaks())
    print(gg)
    png(paste0("results/inc_", r, "_by_disease.png"))
    print(gg)
    dev.off()
}
```

## Incremental by year by region
```{r, echo = F}
region_y_dt <- dt[, .(deaths_averted = sum(incremental)), by = .(region, year)]


gg <- ggplot(region_y_dt[year %in% 2021:2030], aes(x = year, y = deaths_averted / 1e6 , fill = region)) +
  geom_area(color = "white", alpha = 0.8) +
  scale_fill_manual(values = my_colors3, name = "Region") +
  theme_bw() + xlab("Year") + ylab("Deaths averted (in millions)") +
  ggtitle("Deaths averted by region") +
  scale_x_continuous(breaks = scales::pretty_breaks())
print(gg)
png("results/inc_by_region.png")
print(gg)
dev.off()
```

## By disease mean annual incremental deaths averted with min and max annual
```{r, echo = F}
ann_d_dt <- dt[, .(incremental = sum(incremental)), by = .(year, disease)]
d_summ <- ann_d_dt[, .(mean = mean(incremental), min = min(incremental), max = max(incremental)), by = disease]
ann_dt <- dt[, .(incremental = sum(incremental)), by = year]
t_summ <- ann_dt[, .(mean = mean(incremental), min = min(incremental), max = max(incremental))]
t_summ[, disease := "Total"]
d <- rbind(d_summ, t_summ)
gg <- ggplot(d_summ, aes(x = disease, y = mean / 1e3)) + geom_point() +
  geom_errorbar(aes(ymin = min / 1e3, ymax = max / 1e3), width = 0.2) +
  theme_bw() + xlab("Disease") + ylab("Annual incremental deaths averted (in thousands)") +
  ggtitle("Mean and range of annual incremental deaths averted by disease")
print(gg)
png("results/inc_mean_range_disease.png")
print(gg)
dev.off()
kable(d)
```

# Income grouping
```{r}
dt <- fread("outputs/detailed_results_income.csv")
dt[, income_group := factor(income_group, levels = c("Low income", "Lower middle income", "Upper middle income", "High income"))]

```

## Total by disease for each income group

```{r, echo = F}
income_group_d_dt <- dt[, .(deaths_averted = sum(total)), by = .(income_group, year, disease)]
full_dt <- data.table(expand.grid(
    income_group = unique(income_group_d_dt$income_group),
    year = unique(income_group_d_dt$year),
    disease = unique(income_group_d_dt$disease)
))
income_group_d_dt <- merge(income_group_d_dt, full_dt, all.y = T)
income_group_d_dt[is.na(deaths_averted), deaths_averted := 0]

gg <- ggplot(income_group_d_dt[year %in% 2021:2030], aes(x = year, y = deaths_averted / 1e6 , fill = disease)) +
geom_bar(position = "stack", stat = "identity") +
scale_fill_manual(values = my_colors1, name = "Disease") +
theme_bw() + xlab("Year") + ylab("Deaths averted (in millions)") +
facet_wrap(~income_group) +
scale_x_continuous(breaks = scales::pretty_breaks())
print(gg)
png(paste0("results/income_by_disease.png"))
print(gg)
dev.off()

gg <- ggplot(income_group_d_dt[year %in% 2021:2030], aes(x = year, y = deaths_averted / 1e6 , fill = disease)) +
geom_bar(position = "stack", stat = "identity") +
scale_fill_manual(values = my_colors1, name = "Disease") +
theme_bw() + xlab("Year") + ylab("Deaths averted (in millions)") +
facet_wrap(~fct_rev(income_group)) +
scale_x_continuous(breaks = scales::pretty_breaks())
print(gg)
png(paste0("results/income_by_disease_rev.png"))
print(gg)
dev.off()

```

## Total by year by income
```{r, echo = F}
income_group_y_dt <- dt[, .(deaths_averted = sum(total)), by = .(income_group, year)]

gg <- ggplot(income_group_y_dt[year %in% 2021:2030], aes(x = year, y = deaths_averted / 1e6 , fill = income_group)) +
  geom_bar(position = "stack", stat = "identity") +
  scale_fill_manual(values = my_colors3, name = "Income group") +
  theme_bw() + xlab("Year") + ylab("Deaths averted (in millions)") +
  ggtitle("Deaths averted by income group") +
  scale_x_continuous(breaks = scales::pretty_breaks())
print(gg)
png("results/total_by_income.png")
print(gg)
dev.off()

gg <- ggplot(income_group_y_dt[year %in% 2021:2030], aes(x = year, y = deaths_averted / 1e6 , fill = fct_rev(income_group))) +
  geom_bar(position = "stack", stat = "identity") +
  scale_fill_manual(values = my_colors3, name = "Income group") +
  theme_bw() + xlab("Year") + ylab("Deaths averted (in millions)") +
  ggtitle("Deaths averted by income group") +
  scale_x_continuous(breaks = scales::pretty_breaks())
print(gg)
png("results/total_by_income_rev.png")
print(gg)
dev.off()

cast_dt <- dcast(income_group_y_dt[year %in% 2021:2030], income_group ~ year)
write.csv(cast_dt, "results/total_by_income_table.csv", row.names = F)
```


## Incremental by disease for each income group

```{r, echo = F}
income_group_d_dt <- dt[, .(deaths_averted = sum(incremental)), by = .(income_group, year, disease)]
full_dt <- data.table(expand.grid(
    income_group = unique(income_group_d_dt$income_group),
    year = unique(income_group_d_dt$year),
    disease = unique(income_group_d_dt$disease)
))
income_group_d_dt <- merge(income_group_d_dt, full_dt, all.y = T)
income_group_d_dt[is.na(deaths_averted), deaths_averted := 0]
for(r in unique(income_group_d_dt$income_group)) {
    plot_dt <- income_group_d_dt[income_group == r]
    gg <- ggplot(plot_dt[year %in% 2021:2030], aes(x = year, y = deaths_averted , fill = disease)) +
    geom_area(color = "white", alpha = 0.8) +
    scale_fill_manual(values = my_colors1, name = "Disease") +
    theme_bw() + xlab("Year") + ylab("Deaths averted") +
    ggtitle(r) +
  scale_x_continuous(breaks = scales::pretty_breaks())
    print(gg)
    png(paste0("results/inc_", r, "_by_disease.png"))
    print(gg)
    dev.off()
}
```

## Incremental by year by income
```{r, echo = F}
income_group_y_dt <- dt[, .(deaths_averted = sum(incremental)), by = .(income_group, year)]

gg <- ggplot(income_group_y_dt[year %in% 2021:2030], aes(x = year, y = deaths_averted / 1e6 , fill = income_group)) +
  geom_area(color = "white", alpha = 0.8) +
  scale_fill_manual(values = my_colors3, name = "Income group") +
  theme_bw() + xlab("Year") + ylab("Deaths averted (in millions)") +
  ggtitle("Deaths averted by income group") +
  scale_x_continuous(breaks = scales::pretty_breaks())
print(gg)
png("results/inc_by_income.png")
print(gg)
dev.off()
```